resevol: an R package for spatially explicit models of pesticide resistance

### 2022-07-17 notebook/ms/ms.Rmd
Some work on the pest ecology subsection of the manuscript

---
title: 'resevol: an R package for spatially explicit models of pesticide resistance given evolving pest genomes'
author: |
        | A. Bradley Duthie&#xb9;&#xb2;, et al.
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2]  alexander.duthie@stir.ac.uk,
  Biological and Environmental Sciences 3A149 University of Stirling Stirling, FK9
  4LA, UK'
output:
  bookdown::pdf_document2:
    citation_package: natbib
    fig_caption: yes
    toc: false
  html_document: default
  word_document: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[R]{RESEVOL, AN R PACKAGE}
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: resevol.bib
biblio-style: apalike
---

@@

The full output of `mine_gmatrix` is passed to the `run_farm_sim` function, which initialises and simulates an evolving population of pests on a changing landscape.
In this section, we explain the landscape, pest ecology, and evolving pest traits.

@@

Individual pests can be modelled to have several reproductive systems and life histories.

>>>>

---
title: 'resevol: an R package for spatially explicit models of pesticide resistance
  given evolving pest genomes'
author: |
  | A. Bradley Duthie&#xb9;&#xb2;, et al.
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2]  alexander.duthie@stir.ac.uk, Biological and Environmental Sciences 3A149
  University of Stirling Stirling, FK9 4LA, UK'
output:
  pdf_document:
    toc: no
  html_document: default
  bookdown::pdf_document2:
    citation_package: natbib
    fig_caption: yes
    toc: no
  word_document: default
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[R]{RESEVOL, AN R PACKAGE}
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: resevol.bib
biblio-style: apalike
---

@@

The full output of `mine_gmatrix` is passed to the `run_farm_sim` function, which initialises and simulates an evolving population of pests on a changing landscape for a natural number of time steps (`time_steps`).
In this section, we explain the landscape, pest ecology, and evolving pest traits.

@@

Individual pests can be modelled to have several reproductive systems and life histories.
Pest reproductive system can be specified in the `run_farm_sim` function using the `repro` argument, which accepts arguments "asexual" (haploid), "sexual" (monoecious), and "biparental" (dioecious).
For "sexual" pests, the `selfing` argument specifies if self-fertilisation is (`TRUE`) or is not (`FALSE`) allowed.
At the start of a simulation, pests are initialised in a random location.
Initialised pests are age of zero if `rand_age = FALSE` or a random age from zero to `max_age` if `rand_age = TRUE`.
Following initialisation, a single time step proceeds with landscape change (see above), pest aging, feeding, pesticide consumption, movement, reproduction, mortality, and immigration.
Feeding, pesticide consumption, movement, and reproduction all depend on pest age.
Pests feed and consume pesticide from ages `min_age_feed` to `max_age_feed`, move from ages  `min_age_move` to `max_age_move`, and reproduce from ages `min_age_reproduce` to `max_age_reproduce` (all inclusive).
Food accumulated is lost during aging if `baseline_metabolism > 0` and pest age is within `min_age_metabolism` and `max_age_metabolism`.
The option to set minimum and maximum ages events makes it possible to model pests with much different life histories [e.g., @Sudo2018].

```{r, echo=FALSE, fig.height=1.75, fig.width=6, fig.cap = "Overview of simulated events in the resevol R package. Note that metabolism, feeding, pesticide consumption, movement, and reproduction are all subject to a minimum and maximum pest age. Consequently, simulation order might not reflect the order of events from the perspective of a focal pest (e.g., pests might move from ages 1-2, but only feed from ages 2-4)."}
mbox <- function(x0, x1, y0, y1){
    xx <- seq(from=x0, to=x1, length.out = 100);
    yy <- seq(from=y0, to=y1, length.out = 100);
    xd <- c(rep(x0, 100), xx, rep(x1,100), rev(xx));
    yd <- c(yy, rep(y1,100), rev(yy), rep(y0, 100));
    return(list(x=xd, y=yd));
}
par(mar=c(0,0,0,0));
# ===============================================================
plot(x=0, y=0, type="n", xlim=c(0,100), ylim=c(50,95), xaxt="n", yaxt="n",
     xlab="",ylab="");
ibox <- mbox(x0 = 0,  x1 = 10, y0 = 88, y1 = 72);
polygon(x = ibox$x, y = ibox$y, lwd = 3, border = "black", col = "white");
cbox <- mbox(x0 = 15, x1 = 25, y0 = 88, y1 = 72);
polygon(x = cbox$x, y = cbox$y, lwd = 3, border = "black", col = "white");
ubox <- mbox(x0 = 30, x1 = 40, y0 = 88, y1 = 72);
polygon(x = ubox$x, y = ubox$y, lwd = 3, border = "black", col = "white");
nbox <- mbox(x0 = 45, x1 = 55, y0 = 88, y1 = 72);
polygon(x = nbox$x, y = nbox$y, lwd = 3, border = "black", col = "white");
fbox <- mbox(x0 = 60, x1 = 70, y0 = 88, y1 = 72);
polygon(x = fbox$x, y = fbox$y, lwd = 3, border = "black", col = "white");
tbox <- mbox(x0 = 75, x1 = 85, y0 = 88, y1 = 72);
polygon(x = tbox$x, y = tbox$y, lwd = 3, border = "black", col = "white");
tbox <- mbox(x0 = 90, x1 = 100, y0 = 88, y1 = 72);
polygon(x = tbox$x, y = tbox$y, lwd = 3, border = "black", col = "white");
text(x=5,  y=84, labels="Initialise", col="black", cex = 0.5);
text(x=5,  y=80, labels="pests and", col="black", cex = 0.5);
text(x=5,  y=76, labels="landscape", col="black", cex = 0.5);
text(x=20, y=84, labels="Rotate", col="black", cex = 0.5);
text(x=20, y=80, labels="crops and", col="black", cex = 0.5);
text(x=20, y=76, labels="pesticides", col="black", cex = 0.5);

text(x=35, y=80, labels="Mutation", col="black", cex = 0.5);

text(x=35, y=80, labels="Mutation", col="black", cex = 0.5);
text(x=50, y=82, labels="Fitness", col="black", cex = 0.5);
text(x=50, y=78, labels="evaluation", col="black", cex = 0.5);
text(x=65, y=82, labels="Tournament", col="black", cex = 0.5);
text(x=65, y=78, labels="selection", col="black", cex = 0.5);
text(x=80, y=80, labels="Replacement", col="black", cex = 0.5);
arrows(x0=10, x1=14, y0=80, y1=80, lwd=2, length=0.10);
arrows(x0=25, x1=29, y0=80, y1=80, lwd=2, length=0.10);
arrows(x0=40, x1=44, y0=80, y1=80, lwd=2, length=0.10);
arrows(x0=55, x1=59, y0=80, y1=80, lwd=2, length=0.10);
arrows(x0=70, x1=74, y0=80, y1=80, lwd=2, length=0.10);
arrows(x0=85, x1=89, y0=80, y1=80, lwd=2, length=0.10);
arrows(x0 = 80, x1 = 80, y0 = 72, y1 = 68, lwd = 2, length = 0);
arrows(x0 = 80, x1 = 20, y0 = 65, y1 = 65, lwd = 2, length = 0);
arrows(x0 = 20, x1 = 20, y0 = 65, y1 = 72, lwd = 2, length = 0.05);
text(x=73, y=63, labels="No", col="black", cex = 0.5);
rbox <- mbox(x0 = 75, x1 = 85, y0 = 62, y1 = 68);
polygon(x = rbox$x, y = rbox$y, lwd = 3, border = "black", col = "black");
text(x=80, y=65, labels="Termination?", col="white", cex = 0.5);
fbox <- mbox(x0 = 60, x1 = 70, y0 = 62, y1 = 50);
polygon(x = fbox$x, y = fbox$y, lwd = 3, border = "black", col = "white");
text(x=65, y=57, labels="Individual", col="black", cex = 0.5);
text(x=65, y=53, labels="network", col="black", cex = 0.5);
arrows(x0 = 80, x1 = 80, y0 = 62, y1 = 55, lwd = 2, length = 0);
arrows(x0 = 80, x1 = 70, y0 = 55, y1 = 55, lwd = 2, length = 0.05);
text(x=73, y=53, labels="Yes", col="black", cex = 0.5);
```

Pests feed by

Pesticide is consumed during feeding

Movement is modelled by

Pests reproduce by

Immigration occurs.
